Northern Pikeminnow Predation on Juvenile Chinook Salmon in Deadwater Slough, Salmon River, Idaho and Potential Impacts to Local Populations

### 2021-12-30 analysis/paper/paper.Rmd
TB methods edits

To estimate the total consumption potential (i.e., the number or amount of juvenile Chinook Salmon consumed during the peaks of DSR [fall] and NRR [spring] emigrations) of Northern Pikeminnow in Deadwater Slough, we used the Fish Bioenergetics v4.0 application developed by @Deslauriers2017 and applied in the R statistical software [@RCoreTeam2021]. The daily rate of consumption in grams for an individual Northern Pikeminnow was estimated based on the following inputs: predator and prey energy densities, predator start and end weights, and water temperatures. Predator energy density for Northern Pikeminnow was fixed at 6,703 Joules(J)/g [@Deslauriers2017]. Prey energy densities were fixed at 21,500 J/g for juvenile Chinook Salmon [@Moss2016]. The predator start weight was calculated using a weight-length formula [@Parker1995] and the average length of Northern Pikeminnow caught in Deadwater Slough during our study. The average length of Northern Pikeminnow caught during the fall and spring efforts were `r avg_fall_NPM` and `r avg_spring_NPM` mm, respectively, which calculated to average starting weights of 504.4 and 670.5 g. We assumed no growth in individual Northern Pikeminnow, and so predator start and end weights are equal, i.e., we assumed the population biomass was stable. Mean daily water temperatures were summarized from 15-minute interval temperature readings available March 3, 2013 to June 14, 2021 from a gage station approximately 22 river kilometers downstream of Deadwater Slough near Shoup, Idaho [@USGS2016].

The largest uncertainty among model parameters was the proportion of the Northern Pikeminnow diet consisting of fish (e.g., juvenile Chinook Salmon) versus other food items (e.g., invertebrates). Therefore, we varied the proportion of the diet consisting of fish versus invertebrates to range from 10-90%, in 10% intervals, for both categories. Energy densities for invertebrates was fixed at 3,000 J/g. We estimate the g of (any) fish consumed for each model run. For simplicity, we assume that any fish prey have the same energy densities as juvenile Chinook Salmon and we do not differentiate juvenile Chinook Salmon from other fish prey than may be present in Deadwater slough (e.g., juvenile steelhead, juvenile Sockeye Salmon, Redside Shiner). Separate models were also run for a 77-day fall period from September 15 - November 30 and a 92-day period in the spring from March 1 - May 31 to coincide with peak emigrations of DSR and NRR juveniles from the Lemhi River, the largest Chinook Salmon population in the Upper Salmon MPG. This resulted in 18 separate model runs (9 diet scenarios and 2 seasons). Each model run provided an estimate of the amount (grams) of fish consumed by an individual Northern Pikeminnow during the fall or spring time periods.

<!-- During this time, water temperatures typically exceed the range (0-7$^\circ$C) that would illicit concealment behavior or torpor from juvenile Chinook Salmon. -->

To estimate the total biomass of fish potentially consumed by all Northern Pikeminnow in Deadwater Slough, we then multiplied the grams of fish consumed by an individual Northern Pikeminnow  during each period (fall and spring) by the estimated predator population size for that period. The total biomass was further converted to an estimate of the total number of fish consumed each period by dividing by the average weight of DSR (`r avg_DSR_g` g) and NRR (`r avg_NRR_g` g) emigrants in the Upper Salmon MPG. The weight of an average DSR and NRR emigrants was taken from fish caught at seven rotary screw traps located throughout the Upper Salmon MPG upstream of Deadwater Slough during the fall and spring time periods. This calculation assumes that all fish prey items consumed by Northern Pikeminnow are, on average, the same size as DSR and NRR emigrants.

>>>>

To estimate the total consumption potential (i.e., the number or amount of juvenile Chinook Salmon consumed during the peaks of DSR [fall] and NRR [spring] emigrations) of Northern Pikeminnow in Deadwater Slough, we used the Fish Bioenergetics v4.0 application developed by @Deslauriers2017 and applied in the R statistical software [@RCoreTeam2021]. The daily rate of consumption in grams for an individual Northern Pikeminnow was estimated based on the following inputs: predator and prey energy densities, predator start and end weights, and water temperatures. Predator energy density for Northern Pikeminnow was fixed at 6,703 Joules(J)/g [@Deslauriers2017]. Prey energy densities were fixed at 21,500 J/g for juvenile Chinook Salmon [@Moss2016]. The predator start weight was calculated using a weight-length formula [@Parker1995] and the average length of Northern Pikeminnow caught in Deadwater Slough during our study. The average length of Northern Pikeminnow caught during the fall and spring efforts were `r avg_fall_NPM` and `r avg_spring_NPM` mm, respectively, which calculated to average starting weights of 504.4 and 670.5 g. We assumed no growth in individual Northern Pikeminnow, and so predator start and end weights are equal (i.e., we assumed the population biomass was stable). Mean daily water temperatures were summarized from 15-minute interval temperature readings available March 3, 2013 to June 14, 2021 from a gage station approximately 22 river kilometers downstream of Deadwater Slough near Shoup, Idaho [@USGS2016].

The largest uncertainty among model parameters was the proportion of the Northern Pikeminnow diet consisting of fish versus non-fish prey items (e.g., invertebrates). Therefore, we conducted nine model runs with varying proportions of fish in the diet ranging from 10-90%, in 10% increments. Energy densities for invertebrates was fixed at 3,000 J/g. Because we were unable to differentiate juvenile Chinook Salmon from other fish prey consumed, we assume all fish prey have the same energy densities as juvenile Chinook Salmon. Separate models were also run for a 77-day fall period from September 15 - November 30 and a 92-day period in the spring from March 1 - May 31 to coincide with peak emigrations of DSR and NRR juveniles from the Lemhi River, the largest Chinook Salmon population in the Upper Salmon MPG. This resulted in 18 separate model runs (9 diet scenarios and 2 seasons). Each model run provided an estimate of the amount (grams) of fish consumed by an individual Northern Pikeminnow during the fall or spring time periods.

To estimate the total biomass of Chinook Salmon potentially consumed by all Northern Pikeminnow in Deadwater Slough, we multiplied the grams of fish consumed by an individual Northern Pikeminnow by the estimated predator population size during each period (fall and spring). The estimate of total biomass was converted to the total number of fish consumed each period by dividing by the average weight of DSR (`r avg_DSR_g` g) and NRR (`r avg_NRR_g` g) emigrants in the Upper Salmon MPG. The average weights of DSR and NRR emigrants were taken from fish caught at seven rotary screw traps located throughout the Upper Salmon MPG upstream of Deadwater Slough during the fall and spring time periods. This calculation assumes that all fish prey items consumed by Northern Pikeminnow are, on average, the same size as DSR and NRR emigrants.
